An agent-based model for city networks based on interactions between firms

### 2019-09-25 main.tex
Update on Overleaf.

\citep{taylor2001specification} world city network driven by interactions between firms/behavior of firms

\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial

\citep{pan2017mapping} different types of linkages: e.g. services here

@@

    \item The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
    \end{itemize}
\end{enumerate}

@@

Let consider the simplest synthetic setup possible:
\begin{itemize}
    \item cities are distributed randomly in a uniform and isotropic space. The total width of the space $d_{max}$ fixed to 3000km to be approximatively the size of Europe. The number of cities is fixed as $N=700$ to similarly match the European urban system structure (GHS database \citep{Florczyk2019ghs}).
    \item city sizes are attributed with a scaling law (equivalent to have a Zipf law for population and a scaling law for economic size), following an exponent $\beta_0$ (fixed at $\beta_0 = 1.1$ in most experiments when not specified, approximate value for GDP from the GHS database, see \citep{raimbault:halshs-02284897})
    % from the GHS database : scaling GDP - pop ~ 1.1 for Europe, rank size 0.94 -> 1.034 : take 1.1 to simplify - can be paramterized ?
    \item country boundaries are constructed to obtain an approximately equal coverage of the space for a fixed number of countries, what remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing} (\textit{Note: on this point this may be realistic if staying at a continental level, although we should include realistic size of countries; this paper is surely wrong when changing the dataset used and the scale of analysis, working at the scale of Mega-city regions e.g.}). Therefore, we do a basic spatial k-means clustering with $C = 20$ countries and attribute the country accordingly.
    \item firm sector structure: 
\end{itemize}

@@

@@

>>>>

\citep{taylor2001specification} world city network driven by interactions between firms/behavior of firms
\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial 
\citep{pan2017mapping} different types of linkages: e.g. services here, but not only. All sectors should be taken into account first.

The originality of the present paper is to approach these issues within the system of cities framework \citep{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \citep{pumain1997pour}. We examine the interactions of european cities within firm linkages.

% We compare the model to real data on transnational firm networks defined by ownership links at city-level, by the means of the Bureau Van Dijk's \emph {FAME} database with a graph theory application.

@@

    \item The ``geopolitical/sociocultural'' distance remains abstract and should be parametrized (see setup below) and estimated in the data-driven setting.
    \item \textcolor{red} TODO the path dependency $w_{ij}$ could (or should ?) be decomposed into in/out degree for origin and destination ? (then stronger self-reinforcment effect)
    \end{itemize}
\end{enumerate}

@@

Let consider the simplest synthetic setup possible:
\begin{itemize}
    \item cities are distributed randomly in a uniform and isotropic space. The total width of the space $d_{max}$ fixed to 3000km to be approximatively the size of Europe. The number of cities is fixed as $N=700$ to similarly match the European urban system structure (GHS database \citep{Florczyk2019ghs}).
    \item city sizes are attributed with a scaling law (equivalent to have a Zipf law for population and a scaling law for economic size), following an exponent $\beta_0$ (fixed at $\beta_0 = 1.1$ in most experiments when not specified, approximate value for GDP from the GHS database, see \citep{raimbault:halshs-02284897})
    % from the GHS database : scaling GDP - pop ~ 1.1 for Europe, rank size 0.94 -> 1.034 : take 1.1 to simplify - can be paramterized ?
    \item country boundaries are constructed to obtain an approximately equal coverage of the space for a fixed number of countries, what remains consistent with non-correlated city sizes at the scale of the countries as our cities are randomly distributed in space \citep{simini2019testing} (\textit{Note: on this point this may be realistic if staying at a continental level, although we should include realistic size of countries; this paper is surely wrong when changing the dataset used and the scale of analysis, working at the scale of Mega-city regions e.g.}). Therefore, we do a basic spatial k-means clustering with $C = 20$ countries and attribute the country accordingly.
    \item firm sector structure: 
    \begin{itemize}
        \item random probabilities
        \item uniform: then the term of proximity is cancelled, no interest
        \item function of size: larger cities are more diverse, larger cities have more knowledge-based sectors. Test 1: log-normal with slightly increasing width and shifting mode.
        % mode of log normal is exp(mu - sigma ^ 2) ; variance is [exp(sigma^2) - 1]*exp(2 mu + sigma ^ 2)
        % when the pdf is 1 / (x sigma sqrt(2 pi)) * exp (- (ln x - mu) ^ 2 / 2 sigma^2 )
        % -> for largest city stddev = sqrt(variance) = fields / 2 ; for smallest sigma = 1 / nfields
        %  ==> solving for mu and sigma with linear mode and variance as a function of log(E_i)
        % => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X)-1) - ln(e_i)
        % => mu = sigma^2 + ln(e_i)
        %    ~ somehow dirty and not sure that most of the distrib for highest values is within [0,1]
    \end{itemize}
\end{itemize}

! links depend on the sectors: cobb douglas for sector proximity should depend on sector composition ? or several functions ?

@@

\textit{find some indicators for polarization : when big cities take everything or when local clusters organize}

Network indicators:

@@

Sector structure indicators:

\begin{itemize}
    \item sector diversity (Herfindhal ?) 
\end{itemize}
